---
title: "3_Plots"
author: "Henri Chung"
date: "7/21/2020"
output: html_document
---

```{r setup, include=FALSE}
list.of.packages <- c("tidyverse","tidymodels",  "reshape2", "furrr", "future", "doParallel", 
                      "text2vec", "stringi" ,"gridExtra")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, repos = "https://cloud.r-project.org")
invisible(capture.output(lapply(list.of.packages, library, character.only = TRUE, warn.conflicts = FALSE, quietly = T)))
rm(list = ls())
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
source("../code/nltksoil_functions.R")
results_tune <- readRDS("../data/results_tune.RDS")
results_test <- readRDS("../data/results_test.RDS")
results_train <- readRDS("../data/results_train.RDS")
soil_split <- readRDS("../data/soil_split.RDS")
soil_train <- training(soil_split)
soil_test <- testing(soil_split)
otu_id <- readRDS("../data/emp_deblur_90bp.qc_filtered_otu_taxonomy_list.RDS")
```

### Background

This R Markdown documents plots the results from our model workflow described in 1_model and compiled in 2_results. 

##### variables

+ **.taxa** - (character) what taxa category is best to aggregate OTU counts? (otu, species, genus, family, order, class, phylum)

+ **response** - (character) envo_biome_2

+ **.model** - (character) which model works best? (random forest, xgboost)

+ **.no_components** - (numeric) for embedding methods, what length of embedding works best?

+ **.method** - (character) which method of transforming otu counts performs best? (base, abundance, transformpca, transformword, word, pca)

+ **id** - (character) what fold in the CV was used?

+ **.metrics** - (list) weighted macro-averaged performance metrics 

+ **.notes** - (list) output messages from model run

+ **.predictions** - (table) data responses (ground truth) and model predictions

+ **.null** - (table) data responses (ground truth) and model predictions from null model


### Metadata

##### what do our results look like?

```{r}
head(as_tibble(results_test))
colnames(results_test)
#head(as_tibble(results_validation))
```
### what is the class distribution of responses?
```{r}
#read sample data
sampledata <- read.csv("../data/emp_qiime_mapping_qc_filtered.tsv", sep = "\t") %>% 
  filter(grepl("soil", Description))


#Response class distributions
p0 <- sampledata %>%
  pull(envo_biome_2) %>%
  table() %>% 
  as.data.frame() %>%
  filter(Freq > 0) %>% 
  ggplot(aes(x = ., y = Freq, fill = .)) + geom_bar(stat = "identity") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.caption = element_text(size = 14)) +
  geom_text(aes(label=Freq), position=position_dodge(width=0.9), vjust=-0.25) +
  labs(title = "Response class distributions (Envo_biome_2)",
       subtitle = "soil samples only, n = 1728",
       caption = "what is the distribution of response variables?") +
  ylab("Count") 
p0
ggsave(p0, filename = "../outputs/p0.svg", height = 6, width = 8)
```

### What is the distribution of taxa categories?
```{r}
p1data <-  otu_id %>%
  filter(otuid %in% colnames(soil_train)) %>%
  select(taxonomy) %>%
  separate(taxonomy, c("kingdom", "phylum", "class", "order", "family", "genus", "species"), sep = "; ") %>%
  mutate(otu = paste(species, 1:nrow(.))) %>% 
  apply(MARGIN = 2, function(x)length(unique(x))) %>%
  as.data.frame() %>%
  rename_(Freq = ".") %>%
  mutate(names = rownames(.))

p1a <- p1data %>%
  filter(names == "otu") %>%
  ggplot(aes(x =names, y = Freq), fill = "gray") + geom_bar(stat = "identity", show.legend = FALSE) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.caption = element_text(size = 14)) +
  geom_text(aes(label=Freq), position=position_dodge(width=0.9), vjust=-0.25) +
  labs(title = "OTUs",
       caption = "What is the distribution of taxa labels for unique OTUs?") +
  ylab("Count") 
         
p1b <- p1data %>%
  filter(names != "otu") %>%
  mutate(names = factor(names, levels = c("species", "genus", "family", "order", "class", "phylum", "kingdom"))) %>%
  ggplot(aes(x =names, y = Freq, fill = names)) + geom_bar(stat = "identity") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.caption = element_text(size = 14)) +
  geom_text(aes(label=Freq), position=position_dodge(width=0.9), vjust=-0.25) +
  labs(title = "Labeled Taxa",
       caption = "") +
  ylab("") 

p1 <- grid.arrange(p1a, p1b, nrow = 1, name = "Distribution of Taxa Labels")
ggsave(p1, filename = "../outputs/p1.svg", height = 6, width = 8)
```

### Results

##### How well do models perform against each other using performance metrics? 

```{r}
null <- results_test %>%
  unnest(.nullmetrics) %>%
  filter(.no_components == 50 & is.na(.taxa)) %>% 
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa)))  %>%
  select(c(.metric, .estimator, .estimate)) %>%
  mutate(.method = "null")

p2data <- results_test %>% 
  unnest(.metrics) %>%
  filter(.no_components == 50 & is.na(.taxa)) %>% 
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  select(c(.method, .metric, .estimator, .estimate)) %>%
  rbind(null)


p2a <- p2data %>%
  filter(.metric == "f_meas") %>%
  ggplot(aes(x = .method, y = .estimate, fill = .method)) +
  stat_summary(fun.data = quantiles_95, geom="boxplot") +
  facet_wrap(~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Testing: F1 Measure",
       subtitle = "95% CI",
       caption = "Compare model performance using F1 Score") +
  ylab("") + xlab("")
p2a
ggsave(p2a, filename = "../outputs/p2a.svg", height = 6, width = 8)

p2b <- p2data %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = .method, y = .estimate, fill = .method)) +
  stat_summary(fun.data = quantiles_95, geom="boxplot") +
  facet_wrap(~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.caption = element_text(size = 14)) +
  labs(title = "Testing: All Metrics",
       subtitle = "95% CI",
       caption = "Compare model performance using all metrics") +
  ylab("") + xlab("")
p2b
ggsave(p2b, filename = "../outputs/p2b.svg", height = 6, width = 8)
```
### Taxa label comparison

```{r}
null3 <- results_test %>%
  unnest(.nullmetrics) %>%
  select(-c(.metrics)) %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base" | .method == "abundance") %>%
  mutate(.method = "null") 

p3data <- results_test %>% 
  unnest(.metrics) %>%
  select(-c(.nullmetrics)) %>%
  rbind(null3) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) 

p3a <- p3data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base" | .method == "abundance") %>%
  filter(.metric == "f_meas") %>%
  ggplot(aes(x = .taxa, y = .estimate, fill = .method)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_wrap(~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Testing: F1 Measure",
       subtitle = "95 CI",
       caption = "Compare effect of taxa aggregation on F1 Score for Base/Abundance") +
  ylab("") + xlab("")
p3a
ggsave(p3a, filename = "../outputs/p3a.svg", height = 6, width = 8)

p3b <- p3data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base" | .method == "abundance") %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = .taxa, y = .estimate, fill = .method)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_grid(.method~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.caption = element_text(size = 14)) +
  labs(title = "Testing: All Metrics",
       subtitle = "95 CI",
       caption = "Compare effect of taxa aggregation on all metrics for Base/Abundance") +
  ylab("") + xlab("method")
p3b
ggsave(p3b, filename = "../outputs/p3b.svg", height = 6, width = 8)

p3c <- p3data %>%
  filter(.method == "transformword" | .method == "transformpca") %>%
  filter(.metric == "f_meas") %>%
  ggplot(aes(x = .no_components, y = .estimate, fill = .no_components)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_wrap(~.method) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.caption = element_text(size = 14)) +
  labs(title = "Testing: F1",
       subtitle = "95 CI",
       caption = "Compare effect of length of embeddings on F1 Score for PCA/Word embeddings") +
  ylab("") + xlab("")
p3c
ggsave(p3c, filename = "../outputs/p3c.svg", height = 6, width = 8)

p3d <- p3data %>%
  filter(.method == "transformword" | .method == "transformpca") %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = .no_components, y = .estimate, fill = .no_components)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_grid(.method~.metric) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.caption = element_text(size = 14)) +
  labs(title = "Testing: All Metrics",
       subtitle = "95 CI",
       caption = "Compare effect of length of embeddings on all metrics for PCA/Word embeddings") +
  ylab("") + xlab("")
p3d

ggsave(p3d, filename = "../outputs/p3d.svg", height = 6, width = 8)


#p3e <- grid.arrange(p3a, p3c, nrow = 1)
#ggsave(p3e, filename = "../outputs/p3e.svg", height = 6, width = 8)
  
```

### How well do models perform against each other using performance PR and ROC curves?

no PR or ROC curves for null model because there are no threshold values.

```{r}
p4data <- results_test %>% 
  select(c(id, .taxa, .response, .model, .no_components, .method, .pr_curve)) %>%
  unnest(.pr_curve) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) 

p4a <- p4data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .taxa)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Precision Recall Curve for Base",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by taxa aggregation for Base")
p4a
ggsave(p4a, filename = "../outputs/p4a.svg", height = 6, width = 8)

p4b <-  p4data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "abundance") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .taxa)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Precision Recall Curve for Abundance",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by taxa aggregation for Abundance")
p4b
ggsave(p4b, filename = "../outputs/p4b.svg", height = 6, width = 8)

p4c <-  p4data %>%
  filter(.method == "transformword") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Precision Recall Curve for Word Embeddings",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by embedding length for Word Embeddings")
p4c
ggsave(p4c, filename = "../outputs/p4c.svg", height = 6, width = 8)

p4d <-  p4data %>%
  filter(.method == "transformpca") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Precision Recall Curve for PCA Embeddings",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by embedding length for PCA Embeddings")
p4d
ggsave(p4d, filename = "../outputs/p4d.svg", height = 6, width = 8)

p4e <-  p4data %>%
  filter(.method == "pca" | .method == "word") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .level)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.method) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Precision Recall Curve for PCA/Word",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level for PCA")
p4e
ggsave(p4e, filename = "../outputs/p4e.svg", height = 6, width = 8)
```

```{r}

p5data <- results_test %>% 
  select(c(id, .taxa, .response, .model, .no_components, .method, .roc_curve)) %>%
  unnest(.roc_curve) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) 

p5a <- p5data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .taxa)) +
  geom_path() +
  geom_abline(lty = 3) +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
    
  labs(title = "Testing: Receiver Operating Characteristic Curve for Base",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by taxa aggregation for Base")
p5a
ggsave(p5a, filename = "../outputs/p5a.svg", height = 6, width = 8)

p5b <-  p5data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "abundance") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .taxa)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Receiver Operating Characteristic Curve for Abundance",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by taxa aggregation for Abundance")
p5b
ggsave(p5b, filename = "../outputs/p5b.svg", height = 6, width = 8)

p5c <-  p5data %>%
  filter(.method == "transformword") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Receiver Operating Characteristic Curve for Word Embeddings",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by length of embedding for Word Embeddings")
p5c
ggsave(p5c, filename = "../outputs/p5c.svg", height = 6, width = 8)

p5d <- p5data %>%
  filter(.method == "transformpca") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Receiver Operating Characteristic Curve for Base",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by length of embedding for PCA embeddings")
p5d
ggsave(p5d, filename = "../outputs/p5d.svg", height = 6, width = 8)

p5e <- p5data %>%
  filter(.method == "pca" | .method == "word") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .level)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.method) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Testing: Receiver Operating Characteristic Curve for PCA/Word",
       subtitle = "N = 10",
      caption = "Comparison of PR Curves for each response level PCA/Word")
p5e
ggsave(p5e, filename = "../outputs/p5e.svg", height = 6, width = 8)
```

### How does the length of embeddings affect performance metrics?
```{r}


p6a <- results_tune %>% 
  filter(.no_components == 50) %>% 
  filter(.method == "base") %>%
  unnest(.metrics) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = as.factor(mtry), y = as.factor(min_n), fill = .estimate)) +
  geom_tile() +
  facet_grid(.taxa ~ .metric) +
  scale_fill_gradient(limits= c(0.5, 1), low = "white", high = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(angle = 90, hjust = 1, size = 6),plot.caption = element_text(size = 14)) +
  labs(title = "Sensitity to Hyperparameters (Base)") +
  xlab("mtry") + ylab("min_n")
p6a


p6b <- results_tune %>% 
  filter(.no_components == 50) %>% 
  filter(.method == "abundance") %>%
  unnest(.metrics) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = as.factor(mtry), y = as.factor(min_n), fill = .estimate)) +
  geom_tile() +
  facet_grid(.taxa ~ .metric) +
  scale_fill_gradient(limits= c(0.5, 1), low = "white", high = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(angle = 90, hjust = 1, size = 6),plot.caption = element_text(size = 14)) +
  labs(title = "Sensitity to Hyperparameters (Abundance)") +
  xlab("mtry") + ylab("min_n")
p6b

p6c <- results_tune %>% 
  filter(.method == "transformword") %>%
  unnest(.metrics) %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = as.factor(mtry), y = as.factor(min_n), fill = .estimate)) +
  geom_tile() +
  facet_grid(.no_components ~ .metric) +
  scale_fill_gradient(limits= c(0.5, 1), low = "white", high = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(angle = 90, hjust = 1, size = 6),plot.caption = element_text(size = 14)) +
  labs(title = "Sensitity to Hyperparameters (Transformword)") +
  xlab("mtry") + ylab("min_n")
p6c

p6d <- results_tune %>% 
  filter(.method == "transformpca") %>%
  unnest(.metrics) %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = as.factor(mtry), y = as.factor(min_n), fill = .estimate)) +
   scale_colour_gradient(low = "white", high = "black")+
  geom_tile() +
  facet_grid(.no_components ~ .metric) +
  scale_fill_gradient(limits= c(0.5, 1), low = "white", high = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(angle = 90, hjust = 1, size = 6),plot.caption = element_text(size = 14)) +
  labs(title = "Sensitity to Hyperparameters (Transformpca)") +
  xlab("mtry") + ylab("min_n")
p6d

p6e <- results_tune %>% 
  filter(.method == "pca" | .method == "word") %>%
  unnest(.metrics) %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = as.factor(mtry), y = as.factor(min_n), fill = .estimate)) +
  geom_tile() +
  facet_grid(.method ~ .metric) +
  scale_fill_gradient(limits= c(0.5, 1), low = "white", high = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(angle = 90, hjust = 1, size = 6),plot.caption = element_text(size = 14)) +
  labs(title = "Sensitity to Hyperparameters (PCA and Word)") +
  xlab("mtry") + ylab("min_n")
p6e
```

```{r}
null <- results_train %>%
  unnest(.nullmetrics) %>%
  filter(.no_components == 50 & is.na(.taxa)) %>% 
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa)))  %>%
  select(c(.metric, .estimator, .estimate)) %>%
  mutate(.method = "null")

q2data <- results_train %>% 
  unnest(.metrics) %>%
  filter(.no_components == 50 & is.na(.taxa)) %>% 
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  select(c(.method, .metric, .estimator, .estimate)) %>%
  rbind(null)


q2a <- q2data %>%
  filter(.metric == "f_meas") %>%
  ggplot(aes(x = .method, y = .estimate, fill = .method)) +
  stat_summary(fun.data = quantiles_95, geom="boxplot") +
  facet_wrap(~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Training: F1 Measure",
       subtitle = "95% CI",
       caption = "Compare model performance using F1 Score") +
  ylab("") + xlab("")
q2a
ggsave(q2a, filename = "../outputs/q2a.svg", height = 6, width = 8)

q2b <- q2data %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = .method, y = .estimate, fill = .method)) +
  stat_summary(fun.data = quantiles_95, geom="boxplot") +
  facet_wrap(~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Training: All Metrics",
       subtitle = "95% CI",
       caption = "Compare model performance using all metrics") +
  ylab("") + xlab("")
q2b
ggsave(q2b, filename = "../outputs/q2b.svg", height = 6, width = 8)

null3 <- results_train %>%
  unnest(.nullmetrics) %>%
  select(-c(.metrics)) %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base" | .method == "abundance") %>%
  mutate(.method = "null") 

q3data <- results_train %>% 
  unnest(.metrics) %>%
  select(-c(.nullmetrics, mtry, min_n)) %>%
  rbind(null3) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) 

q3a <- q3data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base" | .method == "abundance") %>%
  filter(.metric == "f_meas") %>%
  ggplot(aes(x = .taxa, y = .estimate, fill = .method)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_wrap(~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Training: F1 Measure",
       subtitle = "95 CI",
       caption = "Compare effect of taxa aggregation on F1 Score for Base/Abundance") +
  ylab("") + xlab("")
q3a
ggsave(q3a, filename = "../outputs/q3a.svg", height = 6, width = 8)

q3b <- q3data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base" | .method == "abundance") %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = .taxa, y = .estimate, fill = .method)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_grid(.method~.metric) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Training: All Metrics",
       subtitle = "95 CI",
       caption = "Compare effect of taxa aggregation on all metrics for Base/Abundance") +
  ylab("") + xlab("method")
q3b
ggsave(q3b, filename = "../outputs/q3b.svg", height = 6, width = 8)

q3c <- q3data %>%
  filter(.method == "transformword" | .method == "transformpca") %>%
  filter(.metric == "f_meas") %>%
  ggplot(aes(x = .no_components, y = .estimate, fill = .no_components)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_wrap(~.method) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Training: F1",
       subtitle = "95 CI",
       caption = "Compare effect of length of embeddings on F1 Score for PCA/Word embeddings") +
  ylab("") + xlab("")
q3c
ggsave(q3c, filename = "../outputs/q3c.svg", height = 6, width = 8)

q3d <- q3data %>%
  filter(.method == "transformword" | .method == "transformpca") %>%
  filter(.metric %in% c("accuracy", "f_meas", "kap", "pr_auc", "roc_auc", "sens","recall", "precision")) %>%
  ggplot(aes(x = .no_components, y = .estimate, fill = .no_components)) +
  geom_boxplot() +
  ylim(0.5, 1) +
  facet_grid(.method~.metric) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
  labs(title = "Training: All Metrics",
       subtitle = "95 CI",
       caption = "Compare effect of length of embeddings on all metrics for PCA/Word embeddings") +
  ylab("") + xlab("")
q3d

ggsave(q3d, filename = "../outputs/q3d.svg", height = 6, width = 8)

q4data <- results_train %>% 
  select(c(id, .taxa, .response, .model, .no_components, .method, .pr_curve)) %>%
  unnest(.pr_curve) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) 

q4a <- q4data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .taxa)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Precision Recall Curve for Base",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by taxa aggregation for Base")
q4a
ggsave(q4a, filename = "../outputs/q4a.svg", height = 6, width = 8)

q4b <-  q4data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "abundance") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .taxa)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Precision Recall Curve for Abundance",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by taxa aggregation for Abundance")
q4b
ggsave(q4b, filename = "../outputs/q4b.svg", height = 6, width = 8)

q4c <-  q4data %>%
  filter(.method == "transformword") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Precision Recall Curve for Word Embeddings",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by embedding length for Word Embeddings")
q4c
ggsave(q4c, filename = "../outputs/q4c.svg", height = 6, width = 8)

q4d <-  q4data %>%
  filter(.method == "transformpca") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Precision Recall Curve for PCA Embeddings",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level by embedding length for PCA Embeddings")
q4d
ggsave(q4d, filename = "../outputs/q4d.svg", height = 6, width = 8)

q4e <-  q4data %>%
  filter(.method == "pca" | .method == "word") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = recall, y = precision, group = id, color = .level)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.method) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Precision Recall Curve for PCA/Word",
       subtitle = "N = 10",
       caption = "Comparison of PR Curves for each response level for PCA")
q4e
ggsave(q4e, filename = "../outputs/q4e.svg", height = 6, width = 8)

q5data <- results_train %>% 
  select(c(id, .taxa, .response, .model, .no_components, .method, .roc_curve)) %>%
  unnest(.roc_curve) %>%
  mutate(.taxa = ifelse(is.na(as.character(.taxa)), "OTU", as.character(.taxa))) %>%
  mutate(.taxa = factor(.taxa, levels = c("OTU", "species", "genus", "family", "order", "class", "phylum"))) 

q5a <- q5data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "base") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .taxa)) +
  geom_path() +
  geom_abline(lty = 3) +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
    
  labs(title = "Training: Receiver Operating Characteristic Curve for Base",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by taxa aggregation for Base")
q5a
ggsave(q5a, filename = "../outputs/q5a.svg", height = 6, width = 8)

q5b <-  q5data %>%
  filter(.no_components == 50) %>% 
  filter(.method == "abundance") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .taxa)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.taxa) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Receiver Operating Characteristic Curve for Abundance",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by taxa aggregation for Abundance")
q5b
ggsave(q5b, filename = "../outputs/q5b.svg", height = 6, width = 8)

q5c <-  q5data %>%
  filter(.method == "transformword") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Receiver Operating Characteristic Curve for Word Embeddings",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by length of embedding for Word Embeddings")
q5c
ggsave(q5c, filename = "../outputs/q5c.svg", height = 6, width = 8)

q5d <- q5data %>%
  filter(.method == "transformpca") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .no_components)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.no_components) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Receiver Operating Characteristic Curve for Base",
       subtitle = "N = 10",
        caption = "Comparison of PR Curves for each response level by length of embedding for PCA embeddings")
q5d
ggsave(q5d, filename = "../outputs/q5d.svg", height = 6, width = 8)

q5e <- q5data %>%
  filter(.method == "pca" | .method == "word") %>%
  mutate(.level = gsub("biome", "", .level)) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, group = id, color = .level)) +
  geom_path() +
  coord_equal() +
  facet_grid(.level~.method) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.caption = element_text(size = 14)) +
   
  labs(title = "Training: Receiver Operating Characteristic Curve for PCA/Word",
       subtitle = "N = 10",
      caption = "Comparison of PR Curves for each response level PCA/Word")
q5e
ggsave(q5e, filename = "../outputs/q5e.svg", height = 6, width = 8)
```